---
title: "Russell 2000 Deltas"
format: html
editor: visual
---

```{r}
#packages
library(RPostgreSQL)
library(DBI)
library(dplyr)
library(tidyr)
library(RPostgres)
library(lubridate)
library(readxl)
library(ggplot2)
library(knitr)
library(kableExtra)
library(DescTools)
library(bizdays)
library(RQuantLib)
library(plm)
library(broom)
library(purrr)
library(PerformanceAnalytics)
library(xts)
library(progress)
library(stringr)
library(latex2exp)
```

## Load In Price Data

Make sure that your price data uses ISIN as the stock identifier to align with the Form N-PORT filings. For analysis in the form NPORT paper we use daily log-returns for each asset.

```{r}
#important features included in this data are the adjusted price, isin, sector, and stock universe indicator (includes R1 and R2 stocks)
#Make sure that the price you are using is adjusted for stock splits

crsp_data <- readRDS('/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/cleaned_crsp_with_universe.RDS')
```

## CRSP Daily Returns

```{r}
crsp_df <- crsp_data %>%
  group_by(ISIN) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(
    daily_log_ret = {
      # convert each stock’s price series to xts with its date index
      price_xts <- xts(adj_prc, order.by = date)
      as.numeric(Return.calculate(price_xts, method = "log"))
    }
  ) %>%
  ungroup()

r2_crsp <- crsp_df %>% filter(Universe == "R2")
```

## Load in NPORT Data

For the following functions to work you must follow the same SQL structure that was provided in the python files stored on GitHub: https://github.com/merclay3/SEC-Form-N-PORT-Data-Analysis

```{r}
#sql pulling function
#this function pulls data based on the stocks you are interested in looking at holdings for
#it pulls all the holdings for a particular stock

nport_pull_isin <- function(isin, date, con){
  
  # Create a progress bar object
  pb <- progress::progress_bar$new(
    format = "[:bar] :percent :eta",
    total = length(isin),
    clear = FALSE,
    width = 60
  )  
  
  df <- lapply(isin,
              function(isin){
                  # Tick the progress bar for each ISIN processed
                 pb$tick()
                
                sqlStatement <- paste0("SELECT nport_holdings.series_id, nport_holdings.isin, nport_holdings.period_of_report, nport_holdings.filing_date, nport_holdings.balance, nport_holdings.units, nport_holdings.curcd FROM nport_holdings WHERE nport_holdings.isin = '", isin, "' AND nport_holdings.period_of_report >=", date, ";")
                dbGetQuery(con, sqlStatement)
                })
  names(df) <- isin
  return(df)
}

#this function pulls data based on the cik
nport_pull_cik <- function(cik, date, con){
  df <- lapply(cik,
              function(cik){
                sqlStatement <- paste0("SELECT nport_holdings.cik, nport_holdings.series_id, nport_holdings.isin, nport_holdings.period_of_report, nport_holdings.filing_date, nport_holdings.balance, nport_holdings.units, nport_holdings.curcd FROM nport_holdings WHERE nport_holdings.cik = '", cik, "' AND nport_holdings.period_of_report >=", date, ";")
                dbGetQuery(con, sqlStatement)
                })
  names(df) <- cik
  return(df)
}

#this function pulls data based on the series id
nport_pull_series_id <- function(series_id, date, con){
  df <- lapply(series_id,
              function(series_id){
                sqlStatement <- paste0("SELECT nport_holdings.cik, nport_holdings.series_id, nport_holdings.isin, nport_holdings.period_of_report, nport_holdings.filing_date, nport_holdings.balance, nport_holdings.units, nport_holdings.curcd FROM nport_holdings WHERE nport_holdings.series_id = '", series_id, "' AND nport_holdings.period_of_report >=", date, ";")
                dbGetQuery(con, sqlStatement)
                })
  names(df) <- series_id
  return(df)
}

```

```{r}
#function inputs
con <- dbConnect(Postgres(), 
                    dbname = "dbn",
                    host = "localhost",
                    port = 5434,
                    user = "nport",
                    password = "example2")

date <- "'2020-03-01'"
r2_isins <- unique(r2_crsp$ISIN)
```

```{r}
full_nport_pull <- nport_pull_isin(r2_isins, date, con)
```

## Create a List of Holding Deltas

The following function creates a list that contains a dataframe for each R2000 asset. The data frames are labeled by ISIN and contain the holding information for every fund (identified by series_id) holding the assets from March 2020 to December 2024. These data are the cleaned version of what was pulled directly from SQL and the function calculates the notional change in holdings for each fund-asset pair.

```{r}
get_delta_holdings <- function(crsp_data, holdings_data) {
  # Load the calendar for business day calculations
  load_quantlib_calendars("UnitedStates/NYSE", from = "2019-01-01", to = "2025-12-31")
  # Create a progress bar object
  pb <- progress::progress_bar$new(
    format = "[:bar] :percent :eta",
    total = length(holdings_data),
    clear = FALSE,
    width = 60
  )  

  # Filter out empty data frames from the holdings list
  holdings_list <- holdings_data[lapply(holdings_data, nrow) != 0]

  # Process each data frame in the list
  deltas_list <- lapply(holdings_list, function(df) {
    # Tick the progress bar for each ISIN processed
    pb$tick()
    
    # Data cleaning and filtering
    df <- df %>%
      filter(units == "NS", balance >= 0, curcd == "USD", !is.na(series_id))

    # Get previous business day and a public date
    df$adj_date <- adjust.previous(df$period_of_report, 
                                   "QuantLib/UnitedStates/NYSE")
    

    # Join with CRSP data to get publically available and current prices
    df <- df %>%
      left_join(crsp_data %>% select(ISIN, date, adj_prc, SECTORS, CFACPR, Universe),
                by = c("isin" = "ISIN", "adj_date" = "date")) %>%
      rename(adj_period_of_report = adj_date)
    
    #group together balances
    df <- df %>%
      group_by(series_id, isin, adj_period_of_report) %>%
      # keep only the latest filing
      filter(filing_date == max(filing_date, na.rm = TRUE)) %>%
      # sum balance within that filing
      mutate(balance = sum(balance, na.rm = TRUE)) %>%
      # remove duplicate rows created by the mutate
      distinct(series_id, isin, adj_period_of_report, filing_date, .keep_all = TRUE) %>%
      ungroup()
    
    df <- df %>%
      mutate(
        quarter_indicator = case_when(
          month(adj_period_of_report) %in% c(1, 4, 7, 10) ~ 1,
          month(adj_period_of_report) %in% c(2, 5, 8, 11) ~ 2,
          month(adj_period_of_report) %in% c(3, 6, 9, 12) ~ 3,
          TRUE ~ NA_real_ # For any other cases
          )
      )
      
    # Calculate adjusted balance and total price
    df$adj_balance <- df$balance * df$CFACPR
    df$notional_prc <- df$adj_prc * df$adj_balance
      
    #calculate deltas while checking for valid lags
    df <- df %>%
      mutate(
        year = lubridate::year(adj_period_of_report),
        quarter = lubridate::quarter(adj_period_of_report),
        q_index = year * 4 + quarter
      ) %>%
      group_by(series_id) %>%
      arrange(adj_period_of_report) %>%
      ungroup()
  
  #create new rows for missing lags
    df <- df %>%
      group_by(series_id, isin) %>%
      arrange(q_index, .by_group = TRUE) %>%
      # store first and last observed quarter
      mutate(
        first_q = min(q_index),
        last_q  = max(q_index)
      ) %>%
      # complete ONLY inside observed range
      complete(
        q_index = seq(first_q[1], last_q[1], by = 1)
      ) %>%
      # fill balances only for internally missing quarters
      mutate(
        adj_balance = replace_na(adj_balance, 0),
        notional_prc = replace_na(notional_prc, 0)
      ) %>%
      arrange(q_index, .by_group = TRUE) %>%
      # now difference normally
      mutate(
        bal_delta = adj_balance - dplyr::lag(adj_balance),
        notional_prc_delta = notional_prc - dplyr::lag(notional_prc)
      ) %>%
      ungroup()
    
    #fill new rows with missing values
    df <- df %>%
      group_by(series_id) %>%
      mutate(
        quarter_indicator = first(quarter_indicator[!is.na(quarter_indicator)])
      ) %>%
      ungroup()
    
    df <- df %>%
      arrange(isin, q_index) %>%
      # fill within quarter-security groups
      group_by(isin, quarter_indicator, q_index) %>%
      fill(
        adj_period_of_report,
        quarter_indicator,
        adj_prc,
        Universe,
        SECTORS,
        .direction = "downup"
      ) %>%
      ungroup() %>%
      select(series_id, isin, adj_period_of_report, adj_prc,quarter_indicator, 
             adj_balance, notional_prc, bal_delta, notional_prc_delta)

    return(df)
  })

  # Set the names for the output list
  names(deltas_list) <- names(holdings_list)

  return(deltas_list)
}
```

```{r}
full_nport_deltas <- get_delta_holdings(crsp_data = crsp_df,
                                 holdings_data = full_nport_pull)
```
