---
title: "paper revisions 2"
format: html
editor: visual
---

```{r}
#| label: load-packages
#| include: false

# packages_to_load <- c("RPostgreSQL", "DBI", "dplyr", "tidyverse", "RPostgres",
#                    "lubridate", "readxl", "ggplot2", "knitr", "kableExtra",
#                    "DescTools", "bizdays", "RQuantLib", "plm", "broom", 
#                    "purrr", "PerformanceAnalytics", "xts", "MuMIn", "tidyr",
#                    "progress", "stringr")
# lapply(packages_to_load, library, character.only = TRUE)

library(RPostgreSQL)
library(DBI)
library(dplyr)
library(tidyverse)
library(RPostgres)
library(lubridate)
library(readxl)
library(ggplot2)
library(knitr)
library(kableExtra)
library(DescTools)
library(bizdays)
library(RQuantLib)
library(plm)
library(broom)
library(purrr)
library(PerformanceAnalytics)
library(xts)
library(MuMIn)
library(tidyr)
library(progress)
library(stringr)
library(latex2exp)
library(forecast)
library(nlme)
```

## Data

```{r}
#read in data (this is NOT the same as the original paper results)
delta_df <- readRDS('/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/adj_balance_aggregate_deltas.RDS')

# # outlier removal
# # Calculate mean and sd
# mean_ret <- mean(delta_df$quarterly_return, na.rm = TRUE)
# sd_ret   <- sd(delta_df$quarterly_return, na.rm = TRUE)
# 
# # Define bounds
# lower_bound <- mean_ret - 6 * sd_ret
# upper_bound <- mean_ret + 6 * sd_ret
# 
# # Select outlier ISINs
# outlier_isins <- unique(delta_df$isin[
#   delta_df$quarterly_return < lower_bound |
#   delta_df$quarterly_return> upper_bound
# ])
# 
# length(outlier_isins)
# outlier_isins
# 
# delta_df_cleaned <- delta_df %>%
#   filter(!isin %in% outlier_isins)
```

```{r}
#crsp_data
crsp_data <- readRDS('/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/cleaned_crsp_with_universe.RDS')

crsp_df <- crsp_data %>%
  group_by(ISIN) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(
    daily_log_ret = {
      # convert each stock’s price series to xts with its date index
      price_xts <- xts(adj_prc, order.by = date)
      as.numeric(Return.calculate(price_xts, method = "log"))
    }
  ) %>%
  ungroup()

r2_crsp <- crsp_df %>% filter(Universe == "R2")
```

## Functions

```{r}
future_returns <- function(delta_df, crsp_df, b){
  
  new_date <- paste0("adj_date_b", b)
  new_prc <- paste0("prc_b", b)
  new_returns <- paste0("log_ret_b", b)
  
  delta_df[[new_date]] <- adjust.next(delta_df$adj_period_of_report + b, 
                               "QuantLib/UnitedStates/NYSE")    
  
  
  delta_df <- delta_df %>%
        left_join(crsp_df %>% 
                    select(ISIN, date, adj_prc) %>%
                    rename(!!new_prc := adj_prc),
                  by = c("isin" = "ISIN", setNames("date", new_date)))
        
  
  delta_df <- delta_df %>%
      arrange(isin, adj_period_of_report) %>%
      group_by(isin, quarter_indicator) %>%
      group_split() %>%
      map_df(function(group_df) {
        if (nrow(group_df) > 1) {
          # Convert to xts object
          xts_prices <- xts(group_df[[new_prc]], order.by = group_df[[new_date]])
          
          # Calculate log price returns
          xts_returns <- Return.calculate(xts_prices, method = "log")
          
          # Merge back into original group_df
          group_df[[new_returns]] <- coredata(xts_returns)
        } else {
          group_df[[new_returns]] <- NA
        }
        
        return(group_df)
      })
  
    delta_df[[new_returns]] <- as.numeric(delta_df[[new_returns]])
    
    return(delta_df)
}
```

```{r}
create_pdata_rolling <- function(df, num_panels, dependent_var, independent_var, quarter, index, window_size){
  
  delta_df <- df %>% filter(quarter_indicator == quarter)
  delta_df <- delta_df %>%
    drop_na(all_of(dependent_var), all_of(independent_var))
  
  delta_df <- delta_df %>%
    group_by(isin) %>%
    filter(n() == num_panels) %>%
    arrange(adj_period_of_report) %>%
    ungroup()
  
  # CHANGED: derive split_dates automatically from the sorted unique dates in the data
  all_dates <- sort(unique(delta_df$adj_period_of_report))
  
  # CHANGED: test dates start after window_size periods so there's always a full training window
  split_dates <- all_dates[(window_size + 1):length(all_dates)]
  
  # CHANGED: loop over split_dates, building a fixed-size rolling training window each iteration
  splits <- lapply(split_dates, function(split_date) {
    
    # CHANGED: find the window_size dates immediately before split_date
    prior_dates <- all_dates[all_dates < split_date]
    train_dates <- tail(prior_dates, window_size)   # CHANGED: rolling — only the last window_size periods
    
    train <- delta_df %>%
      filter(adj_period_of_report %in% train_dates)  # CHANGED: filter to exact rolling window dates
    
    test <- delta_df %>%
      filter(adj_period_of_report == split_date)     # UNCHANGED: same filter logic
    
    # UNCHANGED: skip this window if train or test is empty
    if (nrow(train) == 0 || nrow(test) == 0) {
      warning(paste("Skipping split_date", split_date, "- empty train or test set"))
      return(NULL)
    }
    
    train_pdata <- pdata.frame(train, index = index)  # UNCHANGED
    test_pdata  <- pdata.frame(test,  index = index)  # UNCHANGED
    
    list(train = train_pdata, test = test_pdata)      # UNCHANGED: same output structure per window
  })
  
  # UNCHANGED: name each split by its date and drop any NULLs from skipped windows
  names(splits) <- as.character(split_dates)
  splits <- Filter(Negate(is.null), splits)
  
  return(splits)
}
```

```{r}
create_pdata_expanding <- function(df, num_panels, dependent_var, independent_var, quarter, index, split_dates){
  
  delta_df <- df %>% filter(quarter_indicator == quarter)
  delta_df <- delta_df %>%
    drop_na(all_of(dependent_var), all_of(independent_var))
  
  delta_df <- delta_df %>%
    group_by(isin) %>%
    filter(n() == num_panels) %>%
    arrange(adj_period_of_report) %>%
    ungroup()
  
  # CHANGED: loop over split_dates, building an expanding training window each iteration
  splits <- lapply(split_dates, function(split_date) {
    
    train <- delta_df %>%
      filter(adj_period_of_report < split_date)   # UNCHANGED: same filter logic
    
    test <- delta_df %>%
      filter(adj_period_of_report == split_date)  # UNCHANGED: same filter logic
    
    # CHANGED: skip this window if train or test is empty (can happen at edges of date range)
    if (nrow(train) == 0 || nrow(test) == 0) {
      warning(paste("Skipping split_date", split_date, "- empty train or test set"))
      return(NULL)
    }
    
    train_pdata <- pdata.frame(train, index = index)  # UNCHANGED
    test_pdata  <- pdata.frame(test,  index = index)  # UNCHANGED
    
    list(train = train_pdata, test = test_pdata)      # UNCHANGED: same output structure per window
  })
  
  # CHANGED: name each split by its date and drop any NULLs from skipped windows
  names(splits) <- as.character(split_dates)
  splits <- Filter(Negate(is.null), splits)
  
  return(splits)
}
```

```{r}
panel_analysis1 <- function(train_pdata, test_pdata, independent_var, dependent_var,
                           random_effect = TRUE,
                           title1 = "Predicted vs. Actual", 
                           xlab = "Predicted", ylab = "Actual",
                           legend_title = "Estimated Return Classification",
                           title2 = "Out of Sample Estimation Accuracy by Sector"){
  
  #Panel Formula
  formula <- as.formula(paste(dependent_var, "~", independent_var))
  
  # Fixed Effects Model
  fe_model <- plm(formula, data = train_pdata, model = "within")
  
  # Random Effects Model
  re_model <- plm(formula, data = train_pdata, model = "random")
  
  if(random_effect == TRUE){
    final_model <- re_model
  }else{
    final_model <- fe_model
  }
  
  # Hausman Test to choose between FE and RE
  hausman_test <- phtest(fe_model, re_model)
  print(hausman_test)

  if (as.numeric(hausman_test$p.value) < 0.05) {
    #final_model <- fe_model
    print("Based on the Hausman test, a fixed effects model was chosen as the final model.")
  } else {
    #final_model <- re_model
    print("Based on the Hausman test, a random effects model was chosen as the final model.")
  }

  test_pdata <- test_pdata %>%
    mutate(predicted_value = predict(final_model, newdata = test_pdata))
  
  
  #Creating Estimation Table and Color Coded Plot
  df <- test_pdata %>% 
    select(
      actual = dependent_var,
      predicted = predicted_value
    )
  
  # Define band width around y = x
  delta <- .25*sd(df$actual)  # Adjust as needed
  
  # Classify points
  df$category <- NA
  
  # Quadrant 1: Correct Positive Prediction
  df$category[df$predicted > 0 & df$actual > df$predicted - delta] <- "Correct Positive"
  
  # Quadrant 2: Incorrect (False Negative)
  df$category[df$predicted < 0 & df$actual > df$predicted + delta] <- "False Negative"
  
  # Quadrant 3: Correct Negative Prediction
  df$category[df$predicted < 0 & df$actual < df$predicted + delta] <- "Correct Negative"
  
  # Quadrant 4: Incorrect (False Positive)
  df$category[df$predicted > 0 & df$actual < df$predicted - delta] <- "False Positive"
  
  # Convert category to factor for ordered legend
  df$category <- factor(df$category)
  
  # Plot with color coding
  plot <- ggplot(df, aes(x = predicted, y = actual, color = category)) +
      geom_point(alpha = 0.7) +  # Adjust transparency for better visualization
      geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed") +  # y = x line
        geom_abline(slope = 1, intercept = delta, color = "blue") +
      geom_abline(slope = 1, intercept = -delta, color = "blue") +
      scale_color_manual(values = c("Correct Positive" = "darkblue",
                                    "Correct Negative" = "steelblue2",
                                    "False Positive" = "darkorange2",
                                    "False Negative" = "goldenrod1")) +
      labs(title = str_wrap(title1, width = 60),
           x = xlab,
           y = ylab,
           color = legend_title) +
      theme_minimal()
  
  summary_table <- df %>%
    group_by(category) %>%
    summarise(count = n()) %>%
    mutate(percentage = count / sum(count) * 100)
  
  df2 <- df
  df2$isin <- substr(rownames(df2), 1, 12)
  df2 <- as.data.frame(df2)
  test_df <- as.data.frame(test_pdata)
  
  test_df$isin <- as.character(test_df$isin)
  df2$isin   <- as.character(df2$isin)

  df2 <- left_join(df2, test_df, by = "isin")
  
  summary_table2 <- df2 %>%
    group_by(SECTORS, category) %>%
    summarise(count = n()) %>%
    mutate(percentage = count / sum(count) * 100) %>%
    ungroup()
  #summary_table2
  
  
  sector_plot <- ggplot(summary_table2, aes(x = SECTORS, y = count, fill = category)) +
        geom_bar(stat = "identity", position = "stack") +  # Stacked bar chart
        geom_text(aes(label = paste0(round(percentage, 1), "%")), 
                  position = position_stack(vjust = 0.5), size = 3, color = "cornsilk") +  # Add percentage labels
        labs(x = "Sector", 
             y = "Count", 
             title =  str_wrap(title2, width = 60), 
             fill = legend_title) +
        scale_fill_manual(values = c("Correct Positive" = "darkblue",
                                      "Correct Negative" = "steelblue2",
                                      "False Positive" = "darkorange2",
                                          "False Negative" = "goldenrod1")) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels if needed
  
    # --- ADDED: naive benchmark using training mean ---
    train_mean <- mean(as.data.frame(train_pdata)[[dependent_var]], na.rm = TRUE)
    
    naive_df <- df %>%
      mutate(
        naive_predicted = train_mean,
        naive_category  = case_when(
          naive_predicted > 0 & actual > naive_predicted - delta ~ "Correct Positive",
          naive_predicted < 0 & actual > naive_predicted + delta ~ "False Negative",
          naive_predicted < 0 & actual < naive_predicted + delta ~ "Correct Negative",
          naive_predicted > 0 & actual < naive_predicted - delta ~ "False Positive"
        ),
        naive_category = factor(naive_category)
      )
    
    naive_table <- naive_df %>%
      group_by(naive_category) %>%
      summarise(count = n()) %>%
      mutate(percentage = count / sum(count) * 100) %>%
      rename(category = naive_category)
    # --- END ADDED ---
    
    # UNCHANGED: same output list structure as original
    output <- list(final_model, plot, summary_table, sector_plot, 
                   summary_table2, df, naive_table)
    names(output) <- c("final_model", "estimation_plot", "estimation_table",
                       "sector_barplot", "sector_table", "prediction_data", "naive_table")
  

  # output <- list(final_model, plot, summary_table, sector_plot, summary_table2, df)
  # names(output) <- c("final_model", "estimation_plot", "estimation_table", "sector_barplot", "sector_table", "prediction_data")
  
  return(output)
}
```

```{r}
panel_analysis2 <- function(splits, independent_var, dependent_var,
                                     random_effect = TRUE,
                                     title1 = "Predicted vs. Actual",
                                     xlab = "Predicted", ylab = "Actual",
                                     legend_title = "Estimated Return Classification",
                                     title2 = "Out of Sample Estimation Accuracy by Sector") {
  
  # CHANGED: loop over each named split window
  results <- lapply(names(splits), function(split_date) {
    
    cat("\n===== Running panel_analysis for split date:", split_date, "=====\n")
    
    train_pdata <- splits[[split_date]]$train   # UNCHANGED: same variable names as original
    test_pdata  <- splits[[split_date]]$test    # UNCHANGED
    
    # ---- Everything below is UNCHANGED from original panel_analysis ----
    
    formula <- as.formula(paste(dependent_var, "~", independent_var))
    
    fe_model <- plm(formula, data = train_pdata, model = "within")
    re_model <- plm(formula, data = train_pdata, model = "random")
    
    if(random_effect == TRUE){
      final_model <- re_model
    } else {
      final_model <- fe_model
    }
    
    hausman_test <- phtest(fe_model, re_model)
    print(hausman_test)
    
    if (as.numeric(hausman_test$p.value) < 0.05) {
      print("Based on the Hausman test, a fixed effects model was chosen as the final model.")
    } else {
      print("Based on the Hausman test, a random effects model was chosen as the final model.")
    }
    
    test_pdata <- test_pdata %>%
      mutate(predicted_value = predict(final_model, newdata = test_pdata))
    
    df <- test_pdata %>%
      select(
        actual = dependent_var,
        predicted = predicted_value
      )
    
    delta <- .25 * sd(df$actual)
    
    df$category <- NA
    df$category[df$predicted > 0 & df$actual > df$predicted - delta] <- "Correct Positive"
    df$category[df$predicted < 0 & df$actual > df$predicted + delta] <- "False Negative"
    df$category[df$predicted < 0 & df$actual < df$predicted + delta] <- "Correct Negative"
    df$category[df$predicted > 0 & df$actual < df$predicted - delta] <- "False Positive"
    df$category <- factor(df$category)
    
    plot <- ggplot(df, aes(x = predicted, y = actual, color = category)) +
      geom_point(alpha = 0.7) +
      geom_abline(slope = 1, intercept = 0,      color = "black",  linetype = "dashed") +
      geom_abline(slope = 1, intercept =  delta, color = "blue") +
      geom_abline(slope = 1, intercept = -delta, color = "blue") +
      scale_color_manual(values = c("Correct Positive" = "darkblue",
                                    "Correct Negative" = "steelblue2",
                                    "False Positive"   = "darkorange2",
                                    "False Negative"   = "goldenrod1")) +
      labs(title = str_wrap(paste(title1, ":", split_date), width = 60),  # CHANGED: date appended to title
           x = xlab, y = ylab, color = legend_title) +
      theme_minimal()
    
    summary_table <- df %>%
      group_by(category) %>%
      summarise(count = n()) %>%
      mutate(percentage = count / sum(count) * 100)
    
    df2 <- df
    df2$isin <- substr(rownames(df2), 1, 12)
    df2 <- as.data.frame(df2)
    test_df <- as.data.frame(test_pdata)
    
    test_df$isin <- as.character(test_df$isin)
    df2$isin     <- as.character(df2$isin)
    
    df2 <- left_join(df2, test_df, by = "isin")
    
    summary_table2 <- df2 %>%
      group_by(SECTORS, category) %>%
      summarise(count = n()) %>%
      mutate(percentage = count / sum(count) * 100) %>%
      ungroup()
    
    sector_plot <- ggplot(summary_table2, aes(x = SECTORS, y = count, fill = category)) +
      geom_bar(stat = "identity", position = "stack") +
      geom_text(aes(label = paste0(round(percentage, 1), "%")),
                position = position_stack(vjust = 0.5), size = 3, color = "cornsilk") +
      labs(x = "Sector", y = "Count",
           title = str_wrap(paste(title2, ":", split_date), width = 60),  # CHANGED: date appended to title
           fill = legend_title) +
      scale_fill_manual(values = c("Correct Positive" = "darkblue",
                                   "Correct Negative"  = "steelblue2",
                                   "False Positive"    = "darkorange2",
                                   "False Negative"    = "goldenrod1")) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    
    # --- ADDED: naive benchmark using training mean ---
    train_mean <- mean(as.data.frame(train_pdata)[[dependent_var]], na.rm = TRUE)
    
    naive_df <- df %>%
      mutate(
        naive_predicted = train_mean,
        naive_category  = case_when(
          naive_predicted > 0 & actual > naive_predicted - delta ~ "Correct Positive",
          naive_predicted < 0 & actual > naive_predicted + delta ~ "False Negative",
          naive_predicted < 0 & actual < naive_predicted + delta ~ "Correct Negative",
          naive_predicted > 0 & actual < naive_predicted - delta ~ "False Positive"
        ),
        naive_category = factor(naive_category)
      )
    
    naive_table <- naive_df %>%
      group_by(naive_category) %>%
      summarise(count = n()) %>%
      mutate(percentage = count / sum(count) * 100) %>%
      rename(category = naive_category)
    # --- END ADDED ---
    
    # UNCHANGED: same output list structure as original
    output <- list(final_model, plot, summary_table, sector_plot, 
                   summary_table2, df, naive_table)
    names(output) <- c("final_model", "estimation_plot", "estimation_table",
                       "sector_barplot", "sector_table", "prediction_data", "naive_table")
    return(output)
  })
  
  # CHANGED: name the outer list by split date so results[["2023-12-31"]] etc. works intuitively
  names(results) <- names(splits)
  return(results)
}
```

## Iterative Panel Data with Rolling Window

```{r}
load_quantlib_calendars("UnitedStates/NYSE", from = "2019-01-01", to = "2025-12-31")

#looping through differetn day values for the returns series
b_values <- c(1, 5, 10, 20, 30, 60, 65)
results <- list()

for (b in b_values) {
  
  # Create dynamic variable names
  dep_var   <- paste0("log_ret_b", b)
  index_var <- paste0("adj_date_b", b)
  
  # Future returns
  final_df <- future_returns(delta_df, r2_crsp, b = b)
  
  # Create rolling window splits
  splits <- create_pdata_rolling(
    df              = final_df,
    num_panels      = 18,
    dependent_var   = dep_var,
    independent_var = "adjusted_notional_delta",
    quarter         = 3,
    index           = c("isin", index_var),
    window_size     = 8
  )
  
  # Extract the last time point from the rolling splits
  last_date  <- names(splits)[length(splits)]
  last_split <- splits[[last_date]]
  
  # Panel analysis on the last split only
  pdata_output <- panel_analysis1(
    train_pdata     = last_split$train,
    test_pdata      = last_split$test,
    independent_var = "adjusted_notional_delta",
    dependent_var   = dep_var,
    random_effect   = TRUE,
    title1          = paste0("Out of Sample Estimation (b=", b, ", ", last_date, " Russell 2000 Holdings)"),
    title2          = paste0("Out of Sample Estimation Accuracy by Sector (b=", b, ", ", last_date, " Russell 2000 Holdings)"),
    xlab            = "Estimated Return",
    ylab            = "Actual Return",
    legend_title    = "Estimated Return Classification"
  )
  
  # Store results
  results[[paste0("b", b)]] <- pdata_output
  
  # Print outputs
  # cat("\n========== b =", b, "| Last Split Date:", last_date, "==========\n")
  # print(summary(pdata_output$final_model))
  # print(pdata_output$estimation_plot)
  # print(pdata_output$estimation_table)
  # print(pdata_output$sector_barplot)
}
```

```{r}
#this is all trained on the 8 periods leading up to september 2024
b_summary_table <- imap_dfr(results, function(res, split_date) {
  tbl <- res$estimation_table
  model <- res$final_model
  cat_col <- tbl[[1]]
  
  # Create and return a single row for this b_d value
  tibble(
    b = split_date,
    R_Squared = r.squared(model),
    beta = model$coefficients[2],
    z_value = coef(summary(model))[2, "z-value"],
    p_value = coef(summary(model))[2, "Pr(>|z|)"],
    Total_Percent_Correct = sum(tbl$percentage[cat_col %in% c("Correct Negative", "Correct Positive")]),
    True_Positive_Percent  = sum(tbl$percentage[cat_col == "Correct Positive"]),
    True_Negative_Percent  = sum(tbl$percentage[cat_col == "Correct Negative"]),
    False_Negative_Percent = sum(tbl$percentage[cat_col == "False Negative"]),
    False_Positive_Percent = sum(tbl$percentage[cat_col == "False Positive"])
  )
})

# Print the result
print(b_summary_table)
```

## Rolling Window

I forced the rolling window to be 10 time points (a year and a half) because we need \>8 time points for the Granger Causality test.

Should I do the granger causality test on the entire dataframe and not just 10 time points (T=17) because the model performs better with T=8 than T=10 and I think 2 years fo data makes more sense than 2.5

```{r}
df_b1 <- future_returns(delta_df, r2_crsp, b = 1)

splits <- create_pdata_rolling(
    df              = df_b1,
    num_panels      = 18,
    dependent_var   = "log_ret_b1",
    independent_var = "adjusted_notional_delta",
    quarter         = 3,
    index           = c("isin", "adj_period_of_report"),
    window_size     = 8
)
results2 <- panel_analysis2(splits, 
                          independent_var = "adjusted_notional_delta",
                          dependent_var = "log_ret_b1")
```

```{r}
#granger causality test included in plm package for entire data frame T=18
sept_df <- df_b1
sept_df$adj_period_of_report <- as.integer(as.factor(sept_df$adj_period_of_report))
granger_pdata <- pdata.frame(sept_df, index = c("isin", "adj_period_of_report"))
pgrangertest(log_ret_b1 ~ adjusted_notional_delta, data = granger_pdata)
```

Discuss impact of COVID on model performance. The rolling window stops containing 2020 data in March 2023 out of sample estimation

```{r}
final_summary_table <- imap_dfr(results2, function(res, split_date) {
  
  tbl <- res$estimation_table
  naive <- res$naive
  model <- res$final_model
  prediction <- res$prediction_data
  cat_col <- tbl[[1]]
  
  # Create and return a single row for this date
  tibble(
    Date = split_date,
    R_Squared = r.squared(model),
    MAE = mean(abs(prediction$actual - prediction$predicted)),
    RMSE = sqrt(mean((prediction$actual - prediction$predicted)^2)),
    Total_Percent_Correct_Naive = sum(naive$percentage[naive[[1]] %in% c("Correct Negative", "Correct Positive")]),
    Total_Percent_Correct = sum(tbl$percentage[cat_col %in% c("Correct Negative", "Correct Positive")]),
    True_Positive_Percent  = sum(tbl$percentage[cat_col == "Correct Positive"]),
    True_Negative_Percent  = sum(tbl$percentage[cat_col == "Correct Negative"]),
    False_Negative_Percent = sum(tbl$percentage[cat_col == "False Negative"]),
    False_Positive_Percent = sum(tbl$percentage[cat_col == "False Positive"])
  )
})

# Print the result
print(final_summary_table)
```

## September Regression

```{r}
#final model trained with one hold out period
sept_results <- results2[["2024-09-30"]]
sept_resid <- sept_results$prediction_data$actual - sept_results$prediction_data$predicted
sept_rmse <- sqrt(mean((sept_resid)^2))
sept_mae <- mean(abs(sept_resid))

summary(sept_results$final_model)
sept_rmse
sept_mae
sept_results$estimation_table
sept_results$estimation_plot
sept_results$sector_barplot
```

```{r}
ggsave(
  "/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/paper_figures_v2/sept_scatter_b=1_v2.jpg",
  plot = sept_results$estimation_plot,
  width = 8, height = 4, units = "in", dpi = 600
)

ggsave(
  "/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/paper_figures_v2/sept_bar_b=1_v2.jpg",
  plot = sept_results$sector_barplot,
  width = 8, height = 6, units = "in", dpi = 600
)
```

```{r}
#hausman test
sept_pdata <- splits$`2024-09-30`$train

# 2. Define the formula
panel_formula <- log_ret_b1 ~ adjusted_notional_delta

# 3. Estimate the Baseline Models
sept_fe <- plm(panel_formula, data = sept_pdata, model = "within")
sept_re <- plm(panel_formula, data = sept_pdata, model = "random")

# 4. Run the Cluster-Robust Hausman Test
# (method = "aux" uses Wooldridge's regression-based test to allow a robust vcov)
robust_hausman <- phtest(
  panel_formula, 
  data = sept_pdata, 
  method = "aux", 
  vcov = function(x) vcovHC(x, method = "arellano", type = "HC1", cluster = "group")
)

# Print the test results to your console
print(robust_hausman)

```

```{r}
##save september isin list for future reference:
isins <- unique(sept_pdata$isin)
saveRDS(isins, "/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/isins.RDS")
```

## Sector Analysis

```{r}
sept_pdata_train <- splits[["2024-09-30"]]$train
sept_pdata_test <- splits[["2024-09-30"]]$test
sectors <- c(as.character(unique(sept_pdata_train$SECTORS)))
sectors <- sectors[!is.na(sectors)]

train_pdata <- list()
test_pdata <- list()
re_panel_models <- list()
fe_panel_models <- list()

sector_table <- data.frame(matrix(nrow = 11, ncol = 5))
colnames(sector_table) <- c("sector", "n", "significant_FE", "r_squared", "percent_correct")


sector_table$sector <- sectors


for(i in 1:length(sectors)){
  train_pdata[[i]] <- as.data.frame(sept_pdata_train) %>% filter(SECTORS == sectors[i])
  test_pdata[[i]] <- as.data.frame(sept_pdata_test) %>% filter(SECTORS == sectors[i])
  
  fe_panel_models[[i]] <- plm(log_ret_b1 ~ adjusted_notional_delta, 
                data = train_pdata[[i]], 
                model = "within",
                index = c("isin", "adj_period_of_report"))
  
  re_panel_models[[i]] <- plm(log_ret_b1 ~ adjusted_notional_delta, 
                data = train_pdata[[i]], 
                model = "random",
                index = c("isin", "adj_period_of_report"))
  
  print(sectors[i])
  print(phtest(fe_panel_models[[i]], re_panel_models[[i]]))
  
  sector_table$n[i] <- length(unique(train_pdata[[i]]$isin)) 
  sector_table$r_squared[i] <- round(summary(fe_panel_models[[i]])$r.squared[1], 2)
  
  #fixed effects column
  fe_summary <- as.data.frame(summary(fixef(fe_panel_models[[i]])))
  fe_count <- sum(fe_summary$`Pr(>|t|)` < 0.1)
  sector_table$significant_FE[i] <- round(fe_count/sector_table$n[i],4)
  
  #prediction
  #note that you have a different n for the prediction
  test_pdata[[i]] <- pdata.frame(test_pdata[[i]], index = c("isin", "adj_period_of_report"))
  test_pdata[[i]]$predicted_return <- predict(fe_panel_models[[i]], newdata = test_pdata[[i]])
  
  # Define delta for correct classification
  delta <- 0.25 * sd(test_pdata[[i]]$quarterly_return, na.rm = TRUE)

  # Initialize category
  test_pdata[[i]]$category <- NA

  # Apply your logic
  test_pdata[[i]]$category[test_pdata[[i]]$predicted_return > 0 & 
                         test_pdata[[i]]$quarterly_return > test_pdata[[i]]$predicted_return - delta] <- "Correct"

  test_pdata[[i]]$category[test_pdata[[i]]$predicted_return < 0 & 
                         test_pdata[[i]]$quarterly_return > test_pdata[[i]]$predicted_return + delta] <- "Incorrect"

  test_pdata[[i]]$category[test_pdata[[i]]$predicted_return < 0 & 
                         test_pdata[[i]]$quarterly_return < test_pdata[[i]]$predicted_return + delta] <- "Correct"

  test_pdata[[i]]$category[test_pdata[[i]]$predicted_return > 0 & 
                         test_pdata[[i]]$quarterly_return < test_pdata[[i]]$predicted_return - delta] <- "Incorrect"

  # Calculate percent correct
  sector_table$percent_correct[i] <- round(
    mean(test_pdata[[i]]$category == "Correct", na.rm = TRUE), 4
) *100
}
names(test_pdata) <- sectors


print(sector_table)
```

## Appendix A - Additional Quarters

### October Regression

```{r}
pdata1 <- create_pdata_rolling(
    df              = df_b1,
    num_panels      = 18,
    dependent_var   = "log_ret_b1",
    independent_var = "adjusted_notional_delta",
    quarter         = 1,
    index           = c("isin", "adj_period_of_report"),
    window_size     = 8
)

pdata1_output <- panel_analysis2(pdata1, 
                          independent_var = "adjusted_notional_delta",
                          dependent_var = "log_ret_b1")

summary(pdata1_output$`2024-10-31`$final_model)
```

### November Regression

```{r}
pdata2 <- create_pdata_rolling(
    df              = df_b1,
    num_panels      = 18,
    dependent_var   = "log_ret_b1",
    independent_var = "adjusted_notional_delta",
    quarter         = 2,
    index           = c("isin", "adj_period_of_report"),
    window_size     = 8
)

pdata2_output <- panel_analysis2(pdata2, 
                          independent_var = "adjusted_notional_delta",
                          dependent_var = "log_ret_b1")

summary(pdata2_output$`2024-11-29`$final_model)
```

## Appendix C - Sensitivity Analysis

### Standard Error Specification

```{r}

sept_pdata <- splits$`2024-09-30`$train
sept_model <- sept_results$final_model

# 1. Cross-Sectional Dependence Test (Pesaran CD test)
# Null Hypothesis: No cross-sectional dependence
# If p < 0.05, you have cross-sectional dependence.
pcdtest(sept_model, test = "cd")

# 2. Heteroskedasticity Test (Breusch-Pagan test)
# Null Hypothesis: Homoskedasticity
# If p < 0.05, you have heteroskedasticity.
library(lmtest)
bptest(log_ret_b1 ~ adjusted_notional_delta, data = sept_pdata, studentize = TRUE)

# 3. Serial Correlation Test (Breusch-Godfrey/Wooldridge test)
# Null Hypothesis: No serial correlation in idiosyncratic errors
# If p < 0.05, you have serial correlation.
pbgtest(sept_model)

```

```{r}
# vcovHC with cluster = "group" clusters by the first index (asset_id)
se_asset <- vcovHC(sept_model, type = "HC1", cluster = "group")
coeftest(sept_model, vcov = se_asset)
```

```{r}
# vcovDC calculates double-clustered standard errors
se_twoway <- vcovDC(sept_model, type = "HC1")
coeftest(sept_model, vcov = se_twoway)
```

```{r}
# vcovSCC (Spatial Correlation Consistent) calculates Driscoll-Kraay SEs
se_dk <- vcovSCC(sept_model, type = "HC3")
coeftest(sept_model, vcov = se_dk)
```

## Appendix D - LME with AR(1)

```{r}
# Convert pdata to a regular data frame
ar_data <- as.data.frame(sept_pdata_train)

# Ensure identifiers and time are clean types
ar_data$isin <- as.factor(as.character(ar_data$isin))
ar_data$adj_period_of_report <- as.Date(ar_data$adj_period_of_report)

# Should return 17 rows per id if balanced
table(table(ar_data$isin))
```

```{r}
# Random intercepts for each individual and AR(1) error within individual over time
lme_model <- lme(fixed = log_ret_b1 ~ adjusted_notional_delta,  # fixed effects
                 random = ~ 1 | isin,    # random intercept for panel units
                 correlation = corAR1(form = ~ adj_period_of_report | isin),  # AR(1) within panels
                 data = ar_data,
                 method = "REML")  # or method = "ML" for model comparison


summary(lme_model)
#compute marginal and conditional R squared values
r.squaredGLMM(lme_model)
```

```{r}
# Extract residuals
resid_vec <- residuals(lme_model, resType = "normalized")
resid_list <- split(resid_vec, ar_data$isin)
```

```{r}
lag1_autocorr <- sapply(resid_list, function(x) {
  if (length(x) > 1) {
    acf(x, lag.max = 1, plot = FALSE)$acf[2]  # lag 1 autocorrelation
  } else {
    NA  # too short to compute autocorrelation
  }
})

# Remove NAs
lag1_autocorr <- lag1_autocorr[!is.na(lag1_autocorr)]

# lme_hist <- hist(lag1_autocorr, 
#      #breaks = "scott", 
#      main = "Lag-1 Residual Autocorrelation by ISIN", 
#      xlab = "Lag-1 Autocorrelation", 
#      col = "steelblue", 
#      border = "white")

lme_hist <- ggplot(data = NULL, aes(x = lag1_autocorr)) + 
  geom_histogram(
    bins = 20,        
    fill = "steelblue",    
    color = "white"        
  ) +
  labs(
    title = "Lag-1 Residual Autocorrelation by ISIN", # Corresponds to base R 'main'
    x = "Lag-1 Autocorrelation",                      # Corresponds to base R 'xlab'
    y = "Count"                                       # Default y-axis for histograms
  )


print(lme_hist)

boxplot(lag1_autocorr)

summary(lag1_autocorr)

acf_obj <- ACF(lme_model, resType = "normalized")
acf_df <- as.data.frame(acf_obj)
acf_df <- subset(acf_df, lag != 0)

lme_acf <- ggplot(acf_df, aes(x = lag, y = ACF)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_segment(aes(xend = lag, yend = 0)) +
  labs(title = "ACF Plot of LME Model Residuals",
       x = "Lag", y = "ACF") +
  theme_minimal()

print(lme_acf)

```

```{r}
ggsave(
  "/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/paper_figures_v2/lme_hist.jpg",
  plot = lme_hist,
  width = 8, height = 4, units = "in", dpi = 600
)

ggsave(
  "/Users/mlc10/Library/CloudStorage/OneDrive-RiceUniversity/13F Project/NPORT/paper_figures_v2/lme_acf.jpg",
  plot = lme_acf,
  width = 8, height = 4, units = "in", dpi = 600
)
```
